---
title: "BDA_project_group"
format: html
editor: visual
---

```{r}
if (!require(tidybayes)) {
    install.packages("tidybayes")
    library(tidybayes)
}
library(dplyr)
if (!require(brms)) {
    install.packages("brms")
    library(brms)
}
if(!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}
if (!require(metadat)) {
  install.packages("metadat")
  library(metadat)
}
if(!require(cmdstanr)){
    install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
    library(cmdstanr)
}
cmdstan_installed <- function(){
  res <- try(out <- cmdstanr::cmdstan_path(), silent = TRUE)
  !inherits(res, "try-error")
}
if(!cmdstan_installed()){
    install_cmdstan()
}
library(loo)

if (!require(gridExtra)) {
  install.packages("gridExtra")
  library(gridExtra)
}

if (!require(priorsense)) {
  install.packages("priorsense")
  library(priorsense)
}
```

```{r}
setwd("YOUR_DIRECTORY")
load("nhgh.rda")
```

```{r}
nhgh_model <- nhgh |>
  dplyr::select(gh, age, race = re, waist, income) |>
  tidyr::drop_na()

nhgh_model <- nhgh_model |>
  mutate(
    age_z   = as.numeric(scale(age)),
    waist_z = as.numeric(scale(waist))
  )
```

```{r}
summary(nhgh_model$gh)
sd(nhgh_model$gh)
# use a N(5, 2) prior. weakly informative centered around the true data mean, slightly higher SD to account for outliers.
hist(nhgh_model$gh, breaks = 40, col = "gray")
```

```{r}
intercept_formula <- bf(
  gh ~ 1,
  family = "student",
  center = FALSE
)

intercept_priors <- c(
  prior(normal(5, 2), class = "b", coef = "Intercept"),
  prior(exponential(1), class = "sigma")
)

fit_intercept <- brm(
  formula = intercept_formula,
  prior = intercept_priors,
  data = nhgh_model,
  chains = 4,
  iter = 2000,
  warmup = 1000,
)
```

# Pooled model

```{r}
nonhier_formula <- bf(
  gh ~ 1 + age_z + waist_z + race,
  family = student()
)

nonhier_priors <- c(
  prior(normal(5, 2), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "age_z"),
  prior(normal(0, 1), class = "b", coef = "waist_z"),
  prior(normal(0, 1), class = "b"),   # covers race dummy coefficients
  prior(exponential(1), class = "sigma")
)

fit_nonhier <- brm(
  formula = nonhier_formula,
  prior = nonhier_priors,
  data = nhgh_model,
  chains = 4,
  iter = 2000,
  warmup=1000
)
```

## Convergence diagnostics

```{r}
fit_nonhier <- readRDS("fit_nonhier.rds")
summary(fit_nonhier)
```

```{r}
nutsparams_fit_nonhier <- nuts_params(fit_nonhier)

sprintf("Divergence of the model: %i", sum(subset(nutsparams_fit_nonhier, Parameter == "divergent__")$Value))
```

We can see that in our model, the $\hat{R}$ values are all equal to 1.00, meaning that across all parameters, the model has properly converged. This means that the model and the estimates for the parameters are reliable. Both of the bulk and tail Effective Sample Size (ESS) are all larger than 2000, meaning that the posterior distribution has been well explored, and thus, the inference is reliable. As there is no divergence, we can conclude that the HMC algorithm has explored the posterior distribution without any issues.

The estimate for the standardised age parameter is noticeably larger than that of the standardised waist circumference parameter, suggesting that age is a stronger predictor for glycohemoglobin. The reference race group in this case is "Mexican American". The racial group "Non-Hispanic White" has a notably lower predicted level of glycohemoglobin compared to the reference group, while that of the group "Non-Hispanic Black" is notably higher. For the other groups, the estimated levels are similar to the reference group with no clear evidence of difference, as their 95% Credible Interval (CI) includes 0.

# Hierarchical model based on race

```{r}
hier_formula <- bf(
  gh ~ 1 + age_z + waist_z + (1 + age_z + waist_z | race),
  family = "student",
  center = FALSE
)

hier_priors <- c(
  prior(normal(5, 2), class = "b", coef = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "age_z"),
  prior(normal(0, 1), class = "b", coef = "waist_z"),
    
  prior(exponential(1), class = "sd", group = "race", coef = "Intercept"),  # weakly informative? also favors small SD values (gh -> does not have large variation e.g 5.5% -> 6.7%
  prior(exponential(1), class = "sd", group = "race", coef = "age_z"),
  prior(exponential(1), class = "sd", group = "race", coef = "waist_z"),
  prior(lkj(2), class = "cor", group = "race"),
    
  prior(exponential(1), class = "sigma") # weakly informative?
)

fit_hier <- brm(
  formula = hier_formula,
  prior = hier_priors,
  data = nhgh_model,
  chains = 4,
  iter = 2000,
  warmup=1000,
  control = list(adapt_delta = 0.99)
)
```

## Convergence diagnostics

```{r}
fit_hier <- readRDS("fit_hier.rds")
summary(fit_hier)
```

```{r}
nutsparams_fit_hier <- nuts_params(fit_hier)

sprintf("Divergence of the model: %i", sum(subset(nutsparams_fit_hier, Parameter == "divergent__")$Value))
```

We can see that in our model, the $\hat{R}$ values are all \~ 1.0, meaning that across all parameters, the model has properly converged. This means that the model and the estimates for the parameters are reliable. Both of the bulk and tail ESS are all larger than 800, meaning that the posterior distribution has been well explored, and thus, the inference is reliable. As there is 1 divergence, the HMC algorithm has encountered a problematic area when exploring the posterior distribution. However, this does not seem to be a big concern.

```{r}
conditions_fit_hier <- make_conditions(nhgh_model, "race", incl_vars = FALSE)
me_fit_hier <- conditional_effects(fit_hier, conditions = conditions_fit_hier, 
                          re_formula = NULL)
plot(me_fit_hier, ncol = 6, points = TRUE)[1]
```

```{r}
plot(me_fit_hier, ncol = 6, points = TRUE)[2]
```

The model includes varying intercepts by race. From the diagnostics, we have the standard deviation between the races as 0.13, with the 95% CI as \[0.06, 0.34\]. This means there there is a difference between how the model predict glycohemoglobin for each race; however, the difference is not large. From the diagnostics values and the plots above, there is substantial overlapping of the posterior distributions for the varying slopes of age and waist circumference, thus there is no clear evidence that one predictor varies more by race than the other. Across the racial groups, both predictors exhibit rather small slope variability.

# Hierarchical model based on race & income

```{r}
hier_formula_inc_race <- bf(
  gh ~ 1 + age_z + waist_z +
    (1 + age_z + waist_z | race) +
    (1 | income),
  family = "student",
  center = FALSE
)

hier_priors_inc_race <- c(
  prior(normal(5, 2), class = "b", coef = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "age_z"),
  prior(normal(0, 1), class = "b", coef = "waist_z"),
    
  prior(exponential(1), class = "sd", group = "race", coef = "Intercept"),
  prior(exponential(1), class = "sd", group = "race", coef = "age_z"),
  prior(exponential(1), class = "sd", group = "race", coef = "waist_z"),
  prior(lkj(2), class = "cor", group = "race"),
    
  prior(exponential(1), class = "sd", group = "income", coef = "Intercept"),
  prior(exponential(1), class = "sigma")
)

fit_hier_inc_race <- brm(
  formula = hier_formula_inc_race,
  prior = hier_priors_inc_race,
  data = nhgh_model,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  control = list(adapt_delta = 0.99)
)
```

## Convergence diagnostics

```{r}
fit_hier_inc_race <- readRDS("fit_hier_inc_race.rds")
summary(fit_hier_inc_race)
```

```{r}
nutsparams_fit_hier_inc_race <- nuts_params(fit_hier_inc_race)

sprintf("Divergence of the model: %i", sum(subset(nutsparams_fit_hier_inc_race, Parameter == "divergent__")$Value))
```

We can see that in our model, the $\hat{R}$ values are all equal to 1.00, meaning that across all parameters, the model has properly converged. This means that the model and the estimates for the parameters are reliable. Both of the bulk and tail ESS are all larger than 500, meaning that the posterior distribution has been well explored, and thus, the inference is reliable. As there is no divergence, we can conclude that the HMC algorithm has explored the posterior distribution without any issues.

From the model diagnostics, it is clear that there is small intercept variability by race (estimate of the standard deviation = 0.13, 95% CI = \[0.05, 0.34\]) and very little to no intercept variability by income (estimate of the standard deviation = 0.01, 95% CI = \[0.00, 0.03\]). This suggests that income affects glycohemoglobin level to a very limited degree, while race is a more notable factor. Across the racial groups, there is substantial overlapping of the posterior distributions for the varying slopes of age and waist circumference, thus there is no clear evidence that one predictor varies more by race than the other: both predictors exhibit rather small slope variability.

## 

# Hierarchical model based on income

```{r}
hier_formula_inc <- bf(
  gh ~ 1 + age_z + waist_z + (1 | income),
  family = "student",
  center = FALSE
)

hier_priors_inc <- c(
  prior(normal(5, 2), class = "b", coef = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "age_z"),
  prior(normal(0, 1), class = "b", coef = "waist_z"),
      
  prior(exponential(1), class = "sd", group = "income", coef = "Intercept"),
  prior(exponential(1), class = "sigma")
)

fit_hier_inc <- brm(
  formula = hier_formula_inc,
  prior = hier_priors_inc,
  data = nhgh_model,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  control = list(adapt_delta = 0.99)
)
```

## Convergence diagnostics

```{r}
fit_hier_inc <- readRDS("fit_hier_inc.rds")
summary(fit_hier_inc)
```

```{r}
nutsparams_fit_hier_inc <- nuts_params(fit_hier_inc)

sprintf("Divergence of the model: %i", sum(subset(nutsparams_fit_hier_inc, Parameter == "divergent__")$Value))
```

We can see that in our model, the $\hat{R}$ values are all equal to 1.00, meaning that across all parameters, the model has properly converged. This means that the model and the estimates for the parameters are reliable. Both of the bulk and tail ESS are all larger than 1000, meaning that the posterior distribution has been well explored, and thus, the inference is reliable. As there is no divergence, we can conclude that the HMC algorithm has explored the posterior distribution without any issues.

The estimate for the standard deviation of the income intercept is very small at 0.03, with 95% CI at \[0.01, 0.05\], suggesting that different income levels only account for very limited variability for glycohemoglobin prediction.

# Posterior predictive checks

```{r}
#pooled
ppc1 <- pp_check(fit_nonhier) + ggplot2::xlab("Pooled model") + ggplot2::coord_cartesian(xlim = c(0, 15))
ppc11 <- pp_check(fit_nonhier, type="dens_overlay_grouped", group="race") + ggplot2::xlab("Pooled model divided into racial groups") + ggplot2::coord_cartesian(xlim = c(0, 15))

#hierarchical based on race
ppc2 <- pp_check(fit_hier, allow_new_levels=TRUE,sample_new_levels="gaussian") + ggplot2::xlab("Hierarchical model based on race") + ggplot2::coord_cartesian(xlim = c(0, 15))
ppc21 <- pp_check(fit_hier, type="dens_overlay_grouped", group="race",allow_new_levels=TRUE, sample_new_levels="gaussian")+ ggplot2::xlab("Hierarchical model based on race, divided into racial groups") + ggplot2::coord_cartesian(xlim = c(0, 15))

#hierarchical based on race & income
ppc3 <- pp_check(fit_hier_inc_race, allow_new_levels=TRUE,sample_new_levels="gaussian") + ggplot2::xlab("Hierarchical model based on race & income") + ggplot2::coord_cartesian(xlim = c(0, 15))

#hierarchical based on income
ppc4 <- pp_check(fit_hier_inc, allow_new_levels=TRUE,sample_new_levels="gaussian") + ggplot2::xlab("Hierarchical model based on income") + ggplot2::coord_cartesian(xlim = c(0, 15))
ppc41 <- pp_check(fit_hier_inc, type="dens_overlay_grouped", group="income",allow_new_levels=TRUE, sample_new_levels="gaussian")+ ggplot2::xlab("Hierarchical model based on income, divided into income groups") + ggplot2::coord_cartesian(xlim = c(0, 15))
```

```{r}
ppc1
```

```{r}
grid.arrange(ppc1, ppc2, ppc3, ppc4, ncol = 2)
```

```{r}
ppc11
```

```{r}
ppc21
```

```{r}
ppc41
```

We can see that for all 4 models, the predicted data graphs seem to be having a highly similar distribution to the observed data. This high degree of similarity is preserved when examining the distributions for each racial group. When the distribution is examined based on income, there are some income groups whose predictions are less accurate, such as the groups "\< 20000" and "\>= 100000". However, for the group "\< 20000" in particular, this is likely due to its very small sample size of 75 people.

# Other

```{r}
saveRDS(fit_hier, file = "fit_hier.rds")
saveRDS(fit_nonhier, file="fit_nonhier.rds")
saveRDS(fit_hier_inc_race, file="fit_hier_inc_race.rds")
saveRDS(fit_hier_inc, file="fit_hier_inc.rds")
```

```{r}
fit_hier <- readRDS("~/Downloads/fit_hier.rds")
fit_nonhier <- readRDS("~/Downloads/fit_nonhier.rds")
fit_hier_inc_race <- readRDS("~/Downloads/fit_hier_inc_race.rds")
fit_hier_inc <- readRDS("~/Downloads/fit_hier_inc.rds")
```

```{r}
loo_hier <- loo(fit_hier, reloo=TRUE)
loo_nonhier <- loo(fit_nonhier, reloo=TRUE)
# add other loos here
```

```{r}
loo_compare(loo_hier, loo_nonhier)
```

# Model Comparison (PSIS-LOO)

```{r}
loo_intercept <- loo(fit_intercept)
loo_nonhier <- loo(fit_nonhier)
loo_hier_race <- loo(fit_hier)
loo_hier_inc <- loo(fit_hier_inc)
loo_hier_inc_race <- loo(fit_hier_inc_race)
```


```{r}
pareto_k_table(loo_intercept)
pareto_k_table(loo_nonhier)
pareto_k_table(loo_hier_race)
pareto_k_table(loo_hier_inc)
pareto_k_table(loo_hier_inc_race)
```

We get that all Pareto-k estimates for all the five models are < 0.7, meaning that PSIS-LOO is reliable for all models (there is no need for moment matcching or refitting). 

```{r}
# printing individual loo for each model
loo_intercept
loo_nonhier
loo_hier_race
loo_hier_inc
loo_hier_inc_race
```

The raw elpd_loo values show that the hierarchical model with varying intercepts and slopes by race (hier_race) has the highest predictive accuracy (elpd_loo = –4766.6). The model that additionally includes income as a grouping factor (hier_inc_race) has nearly identical predictive performance (elpd_loo = –4767.3), differing by only 0.7 units, which is far smaller than one standard error (SE = 108.4). This indicates that adding income-level random effects does not meaningfully improve out-of-sample prediction.

The non-hierarchical model (nonhier) performs moderately worse (elpd_loo = –4772.3), suggesting that modeling race as a hierarchical grouping structure improves predictive accuracy, although the improvement is modest.

In contrast, the hierarchical model based solely on income (hier_inc) shows a much worse elpd_loo (–4900.6), indicating a large decline in predictive accuracy. This confirms that income alone provides very weak, or even misleading, grouping structure for predicting glycohemoglobin in this dataset.

The elpd_loo for intercept only model is extremely low (-5782.9), indicating that it has the worst predictive accuracy among all the other models. all the other non-hierarchical and hierarchical models have a much higher elpd-loo. 

This conclusion is in line with ppc checks done above.

```{r}
# loo comparison between models
loo_compare(
  loo_intercept,
  loo_nonhier,
  loo_hier_race,
  loo_hier_inc,
  loo_hier_inc_race
)
```

From the output of the PSIS-loo comparison, we see that hierarchical model by race has elpd_diff 0, indicating that it is the best model.

The hierarchical model by race and income has elpd_diff = -0.6 and se_diff = 1.3. Since |-0.6| < 1.3, the difference is far within 1*SE, meaning that this model is statistically indistinguishable from the best model (hier_race) in terms of predictive accuracy.

On the other hand, the poold model has elpd_diff = -5.7 and se_diff = 4.6. Since |-5.7| ≈ 1.2*4.6, this model is somewhat worse than the hierarchical race model, but the evidence is not extrmely strong. The hierarchical structure provides better predictive accuracy, but only modestly.

The hierarchical model by income on the other hand performs much worse, with elpd_diff = -134 and se_diff 18.9, where |-134| >> 2*18.9, which is a dramatic difference. The hierarchical model with income as the only grouping factor performs substantially worse than any of the other models. Income alone provides very little grouping structure for predicting glycohemoglobin.

This loo compparison further confirms that fit_intercept is catastrophically worse (by ~1000 elpd units). This makes sense since this model assumes no effect of age, waist circumferece, rances, or income levels, and hence is a massively oversimplified model, leading to aawful predictive accuracy.

# Sensitivity Analysis (priorsense)
Since the hierarchical model by race and the hierarchical model by race and income proved to be the models with the highest preditive accuracies, we will conduct sensitivity analysis for these two models using priorsense.


```{r}
# Key fixed effects we care about
vars_fixed_main <- c(
  "b_Intercept",
  "b_age_z",
  "b_waist_z"
)

# Key group-level SDs (race & inocme)
vars_group_main <- c(
  "sd_race__Intercept",
  "sd_race__age_z",
  "sd_race__waist_z",
  "sd_income__Intercept"
)

# distributional parameters
vars_dist_main <- c("sigma", "nu")

vars_main <- c(vars_fixed_main, vars_group_main, vars_dist_main)
vars_main
```


```{r}
ps_data_main <- create_priorsense_data(fit_hier_inc_race)

ps_main <- powerscale_sensitivity(
  ps_data_main,
  variable  = vars_main,
  component = c("prior", "likelihood"),
  lower_alpha = 0.99,
  upper_alpha = 1.01,
  sensitivity_threshold = 0.05
)

ps_main
```

From the sensitivity analysis output of the hierarchical by race and income model, we see that only two parameters are above the 0.05 sensitivity threshold:
1. sd_race_intercept is somewhat sensitive to prior: This is an expeted effect since the racce-level sd is estimated from only 5 race groups, which is a very small number. Nevertheless, the magnitude of the sensitivity is mild (~0.06).

2. sd_race_age_z shows potential prior-data confilct: This simply means that group-level slope varies very weakly across races (as we saw in the posterior). Because it is so small and barely identified, the posteriior naturally depends a bit more on the prior. 

All the main fixed effects (b_age_z, b-waist_z, b_intercept) are extremely stable. sd-income-intercept is also prior/likelihood stable, which makes sense becauuse it is very close to 0 and does not do much in the model anyways.

```{r}
vars_fixed_race <- c(
  "b_Intercept",
  "b_age_z",
  "b_waist_z"
)

vars_group_race <- c(
  "sd_race__Intercept",
  "sd_race__age_z",
  "sd_race__waist_z"
)

vars_dist_race <- c("sigma", "nu")

vars_race <- c(vars_fixed_race, vars_group_race, vars_dist_race)
vars_race
```


```{r}
ps_data_race <- create_priorsense_data(fit_hier)

ps_race <- powerscale_sensitivity(
  ps_data_race,
  variable  = vars_race,
  component = c("prior", "likelihood"),
  lower_alpha = 0.99,
  upper_alpha = 1.01,
  sensitivity_threshold = 0.05
)

ps_race
```

In this model show even less sensitivity than the race&income model. Only sd_race_intercept hits the 0.05 threshold. This is again due to having only 5 race levels, which causes unavoidable prior influence. We should also note that sds of random effects are always the most prior-sensitive parameters. 


```{r}
vars_fixed_nonhier <- c(
  "b_Intercept",
  "b_age_z",
  "b_waist_z"
)

vars_race_nonhier <- c(
  "b_raceOtherHispanic",
  "b_raceNonMHispanicWhite",
  "b_raceNonMHispanicBlack",
  "b_raceOtherRaceIncludingMultiMRacial"
)

vars_dist_nonhier <- c("sigma", "nu")

vars_nonhier <- c(vars_fixed_nonhier, vars_race_nonhier, vars_dist_nonhier)
vars_nonhier

```

```{r}
ps_data_nonhier <- create_priorsense_data(fit_nonhier)

ps_nonhier <- powerscale_sensitivity(
  ps_data_nonhier,
  variable  = vars_nonhier,
  component = c("prior", "likelihood"),
  lower_alpha = 0.99,
  upper_alpha = 1.01,
  sensitivity_threshold = 0.05
)

ps_nonhier
```

We an see that for this model the prior sensitivity is zero for all regression coefficients. likelihood sensitivity is around 0.08-0.10 for all parameeters, which is normal for a student-t likelihood. 

Hence, the non-hierarchical model is the most stable of all models, because it has no partially-pooled variance components. The data overwhelm the priors on every parameter.

```{r}
vars_fixed_inc <- c(
  "b_Intercept",
  "b_age_z",
  "b_waist_z"
)

vars_group_inc <- c(
  "sd_income__Intercept"
)

vars_dist_inc <- c("sigma", "nu")

vars_inc <- c(vars_fixed_inc, vars_group_inc, vars_dist_inc)
vars_inc
```

```{r}
ps_data_inc <- create_priorsense_data(fit_hier_inc)

ps_inc <- powerscale_sensitivity(
  ps_data_inc,
  variable  = vars_inc,
  component = c("prior", "likelihood"),
  lower_alpha = 0.99,
  upper_alpha = 1.01,
  sensitivity_threshold = 0.05
)

ps_inc
```


In this model also the fixed effects are extremely stable. sd_income_intercept shows ~0.093 likelihood sensitivity whihc is expected sinec the variance is very small (~0.03), between-income variation is minimal, and group-level SDs with small true value usually have noisier likelihood. 

Income contributes extremely little hierarchical structure, which we already saw in the model summary (sd_income__Intercept ≈ 0.03). The sensitivity analysis confirms that the group effect is weak but stable, and the fixed predictors remain robust.


Baseline intercept model:
```{r}
vars_intercept <- c(
  "b_Intercept",
  "sigma",
  "nu"
)

ps_data_intercept <- create_priorsense_data(fit_intercept)

ps_intercept <- powerscale_sensitivity(
  ps_data_intercept,
  variable  = vars_intercept,
  component = c("prior", "likelihood"),
  lower_alpha = 0.99,
  upper_alpha = 1.01,
  sensitivity_threshold = 0.05
)

ps_intercept
```

We an see that all prior sensitivities are basically zero. Likelihood sensitivity is ~0.09-0.11, which is similar to what we observed in the nonhierarchical and the income only models. This happens because student-t models have a heavy-tailed likelihood; sigma and nu usually show moderate sensitivity; and with only one regeression parameter, the likelihood power-scaling changes model fit slightly. 

Overall, the model is exxtremely stable but has terrible predictive performance.

## Some plots to visualize the sensitivity analysis
prior power-scaling: densities for age and waist effects (fit_hier_inc_race)
```{r}
ps_seq_prior_main <- powerscale_sequence(
  ps_data_main,
  component = "prior"
)

powerscale_plot_dens(
  ps_seq_prior_main,
  variable = c("b_age_z", "b_waist_z")
)
```

prior power scaling: mean/SDs for race and income SDs 
```{r}
powerscale_plot_quantities(
  ps_seq_prior_main,
  variable  = c("sd_race__Intercept", "sd_income__Intercept"),
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
)
```

likelihood power scaling: age effect
```{r}
ps_seq_lik_main <- powerscale_sequence(
  ps_data_main,
  component = "likelihood"
)

powerscale_plot_quantities(
  ps_seq_lik_main,
  variable  = "b_age_z",
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
)
```

hier_race model:
```{r}
ps_seq_prior_race <- powerscale_sequence(
  ps_data_race,
  component = "prior"
)

powerscale_plot_dens(
  ps_seq_prior_race,
  variable = c("b_age_z", "b_waist_z")
)
```


nonhier model:
```{r}
ps_seq_prior_nonhier <- powerscale_sequence(
  ps_data_nonhier,
  component = "prior"
)

powerscale_plot_dens(
  ps_seq_prior_nonhier,
  variable = c("b_age_z", "b_waist_z")
)
```

```{r}
powerscale_plot_quantities(
  ps_seq_prior_nonhier,
  variable  = vars_race_nonhier,
  quantity  = "mean"
)
```

hier_inc model:
```{r}
ps_seq_prior_inc <- powerscale_sequence(
  ps_data_inc,
  component = "prior"
)

powerscale_plot_dens(
  ps_seq_prior_inc,
  variable = c("b_age_z", "b_waist_z")
)
```

```{r}
powerscale_plot_quantities(
  ps_seq_prior_inc,
  variable  = "sd_income__Intercept",
  quantity  = c("mean", "sd")
)
```


Baseline intercept model:
```{r}
ps_seq_prior_intercept <- powerscale_sequence(
  ps_data_intercept,
  component = "prior"
)

powerscale_plot_dens(
  ps_seq_prior_intercept,
  variable = c("b_Intercept")
)
```

```{r}
powerscale_plot_quantities(
  ps_seq_prior_intercept,
  variable = c("sigma", "nu"),
  quantity = c("mean", "sd")
)
```


```{r}
if (!require(patchwork)) {
  install.packages("patchwork")
}
library(patchwork)
library(ggplot2)

bda_theme <- function() {
  theme_minimal(base_size = 11) +
    theme(
      plot.subtitle   = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_line(linewidth = 0.3, linetype = "dashed", colour = "grey85"),
      panel.grid.major.y = element_line(linewidth = 0.3, colour = "grey90"),
      strip.background  = element_rect(fill = "grey95", colour = NA),
      strip.text        = element_text(face = "bold"),
      plot.title        = element_text(face = "bold", size = 12),
      axis.title        = element_text(size = 10),
      legend.position   = "bottom"
    )
}
```


```{r}
# Densities for age/waist
p_main_dens <- powerscale_plot_dens(
  ps_seq_prior_main,
  variable = c("b_age_z", "b_waist_z")
) +
  bda_theme() +
  labs(
    title = "Race + income: fixed effects",
    x = "Value", y = "Density",
    colour = expression(alpha)
  )

# Mean/SD for race & income SDs
p_main_sd <- powerscale_plot_quantities(
  ps_seq_prior_main,
  variable  = c("sd_race__Intercept", "sd_income__Intercept"),
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
) +
  bda_theme() +
  labs(
    title = "Race + income: group-level SDs",
    x = expression(alpha), y = NULL
  )

```

```{r}
p_race_dens <- powerscale_plot_dens(
  ps_seq_prior_race,
  variable = c("b_age_z", "b_waist_z")
) +
  bda_theme() +
  labs(
    title = "Race only: fixed effects",
    x = "Value", y = "Density",
    colour = expression(alpha)
  )

p_race_sd <- powerscale_plot_quantities(
  ps_seq_prior_race,
  variable    = c("sd_race__Intercept", "sd_race__age_z", "sd_race__waist_z"),
  quantity    = c("mean", "sd"),
  div_measure = "cjs_dist"
) +
  bda_theme() +
  labs(
    title = "Race only: group-level SDs",
    x = expression(alpha), y = NULL
  )

```

```{r}
p_nonhier_dens <- powerscale_plot_dens(
  ps_seq_prior_nonhier,
  variable = c("b_age_z", "b_waist_z")
) +
  bda_theme() +
  labs(
    title = "Non-hierarchical: fixed effects",
    x = "Value", y = "Density",
    colour = expression(alpha)
  )

p_nonhier_race <- powerscale_plot_quantities(
  ps_seq_prior_nonhier,
  variable  = vars_race_nonhier,  # your vector of race b_* names
  quantity  = "mean",
  div_measure = "cjs_dist"
) +
  bda_theme() +
  labs(
    title = "Non-hierarchical: race coefficients",
    x = expression(alpha), y = NULL
  )

```

```{r}
p_inc_dens <- powerscale_plot_dens(
  ps_seq_prior_inc,
  variable = c("b_age_z", "b_waist_z")
) +
  bda_theme() +
  labs(
    title = "Income only: fixed effects",
    x = "Value", y = "Density",
    colour = expression(alpha)
  )

p_inc_sd <- powerscale_plot_quantities(
  ps_seq_prior_inc,
  variable  = "sd_income__Intercept",
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
) +
  bda_theme() +
  labs(
    title = "Income only: income SD",
    x = expression(alpha), y = NULL
  )

```


```{r}
p_int_dens <- powerscale_plot_dens(
  ps_seq_prior_intercept,
  variable = "b_Intercept"
) +
  bda_theme() +
  labs(
    title = "Intercept-only: intercept",
    x = "Value", y = "Density",
    colour = expression(alpha)
  )

p_int_quant <- powerscale_plot_quantities(
  ps_seq_prior_intercept,
  variable  = c("sigma", "nu"),
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
) +
  bda_theme() +
  labs(
    title = "Intercept-only: sigma and nu",
    x = expression(alpha), y = NULL
  )

```


```{r}
strip_yaxis <- function(p) {
  p + theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
}
```

```{r}
p_main_sd  <- strip_yaxis(p_main_sd)
p_race_sd  <- strip_yaxis(p_race_sd)
p_inc_sd   <- strip_yaxis(p_inc_sd)
p_nonhier_race <- strip_yaxis(p_nonhier_race)
p_int_quant <- strip_yaxis(p_int_quant)

```


```{r}
wider_left <- function(p) {
  p + theme(
    strip.text.y = element_text(size = 7),       # smaller facet labels
    axis.text.y  = element_text(size = 7),       # smaller numeric labels
    plot.margin  = margin(t = 5.5, r = 5.5, b = 5.5, l = 20)  # more left margin
  )
}

```

```{r}
p_int_quant    <- wider_left(p_int_quant)
p_nonhier_race <- wider_left(p_nonhier_race)
p_inc_sd       <- wider_left(p_inc_sd)
p_race_sd      <- wider_left(p_race_sd)
p_main_sd      <- wider_left(p_main_sd)

```

```{r}
fig_sens_all <- (
  p_int_dens      | p_int_quant      ) /
  (p_nonhier_dens | p_nonhier_race   ) /
  (p_inc_dens     | p_inc_sd         ) /
  (p_race_dens    | p_race_sd        ) /
  (p_main_dens    | p_main_sd        ) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

fig_sens_all

```



```{r}
ggsave("fig_sensitivity_all.pdf", fig_sens_all,
       width = 8, height = 12)  
```


```{r}
theme_sens <- function() {
  theme_minimal(base_size = 18) +
    theme(
      plot.title      = element_text(size = 22, face = "bold", hjust = 0.5),
      strip.text      = element_text(size = 18, face = "bold"),
      axis.title      = element_text(size = 18),
      axis.text       = element_text(size = 16),
      legend.position = "bottom",
      legend.text     = element_text(size = 16),
      legend.title    = element_text(size = 18, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey85"),
      plot.margin      = margin(10, 10, 10, 10)
    )
}

remove_subtitle <- function(p) {
  p + theme(plot.subtitle = element_blank())
}

pretty_dens <- function(p) {
  p +
    theme_sens() +
    remove_subtitle() +
    scale_color_manual(values = c("0.8" = "#1B9E77", "1" = "#D95F02", "1.25" = "#7570B3")) +
    guides(colour = guide_legend(title = expression(alpha)))
}

pretty_quant <- function(p) {
  p +
    theme_sens() +
    remove_subtitle() +
    theme(
      axis.text.y  = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    scale_color_manual(values = c("prior" = "#1B9E77", "likelihood" = "#7570B3")) +
    scale_shape_manual(values = c(16, 17))
}

```


Baseline intercept only model
```{r}
# Prior power-scaling sequences
ps_seq_prior_intercept <- powerscale_sequence(
  ps_data_intercept,
  component = "prior"
)

# Raw plots
p_int_dens_raw <- powerscale_plot_dens(
  ps_seq_prior_intercept,
  variable = "b_Intercept"
)

p_int_quant_raw <- powerscale_plot_quantities(
  ps_seq_prior_intercept,
  variable  = c("sigma", "nu"),
  quantity  = c("mean", "sd"),
  div_measure = "cjs_dist"
)

# Pretty versions
p_int_dens    <- pretty_dens(p_int_dens_raw,   "Intercept-only: intercept")
p_int_quant   <- pretty_quant(p_int_quant_raw, "Intercept-only: σ and ν")

# 2-panel slide figure
fig_sens_intercept <- p_int_dens | p_int_quant +
  plot_annotation(
    title = "Sensitivity analysis — Intercept-only model",
    theme = theme(plot.title = element_text(size = 26, face = "bold", hjust = 0.5))
  )

fig_sens_intercept

ggsave("sens_intercept_slide.pdf", fig_sens_intercept,
       width = 12, height = 6, dpi = 300)

```



```{r}
ps_seq_prior_nonhier <- powerscale_sequence(
  ps_data_nonhier,
  component = "prior"
)

p_nonhier_dens_raw <- powerscale_plot_dens(
  ps_seq_prior_nonhier,
  variable = c("b_age_z", "b_waist_z")
)

# vars_race_nonhier should be your vector of race coefficient names
# e.g. c("b_raceOtherHispanic", "b_raceNonMHispanicWhite", ...)
p_nonhier_race_raw <- powerscale_plot_quantities(
  ps_seq_prior_nonhier,
  variable  = vars_race_nonhier,
  quantity  = "mean",
  div_measure = "cjs_dist"
)

p_nonhier_dens <- pretty_dens(p_nonhier_dens_raw, "Non-hierarchical: fixed effects")
p_nonhier_race <- pretty_quant(p_nonhier_race_raw, "Non-hierarchical: race coefficients")

fig_sens_nonhier <- p_nonhier_dens | p_nonhier_race +
  plot_annotation(
    title = "Sensitivity analysis — Non-hierarchical model",
    theme = theme(plot.title = element_text(size = 26, face = "bold", hjust = 0.5))
  )

fig_sens_nonhier

ggsave("sens_nonhier_slide.pdf", fig_sens_nonhier,
       width = 12, height = 6, dpi = 300)

```